---
title: "Final Assignment"
author: Alex Salo  Frederick Peña Sofia Villamil
format: html
editor: visual
editor_options: 
  chunk_output_type: inline
---

# Final Assignment

### Libraries

```{r}
rm(list = ls())

getwd()
setwd("C:/Users/sofia/Documents/Documentos/Master Computational Social Sciences/Survey Research Methodology II/Final Assignment/surveyMethodologyAssignment-24")
```

```{r}
library(haven)
library(readr)
library(haven)
library(writexl)
library(sjlabelled)
library(tidyverse)
library(corrplot)
library(dplyr)
library(gmodels)
library(scales)
library(rnaturalearth)
library(sf)
library(ggplot2)
library(leaflet)
library(car)
library(caret)

```

### Data

```{r}
data <- read_dta("Data/ZA7575.dta")

```

## Cleaning the data

Cleaning specific variables to determine which are the more important once for our analysis.

**ID**

```{r}
data_cleaning_1 <- data %>%
  select(
         -c(studyno1, studyno2, doi, version, edition,survey))

data_cleaning_2 <- data_cleaning_1 %>% 
  select(serialid, everything())
```

#### Country ID

```{r}
# creating a variable for country name to be more organized
data_cleaning_2$country_name <- NA
data_cleaning_2$country_name[data_cleaning_2$q1_1 == 1] <- 'Belgium'
data_cleaning_2$country_name[data_cleaning_2$q1_2 == 1] <- 'Denmark'
data_cleaning_2$country_name[data_cleaning_2$q1_3 == 1] <- 'Germany'
data_cleaning_2$country_name[data_cleaning_2$q1_4 == 1] <- 'Greece'
data_cleaning_2$country_name[data_cleaning_2$q1_5 == 1] <- 'Spain'
data_cleaning_2$country_name[data_cleaning_2$q1_6 == 1] <- 'France'
data_cleaning_2$country_name[data_cleaning_2$q1_7 == 1] <- 'Ireland'
data_cleaning_2$country_name[data_cleaning_2$q1_8 == 1] <- 'Italy'
data_cleaning_2$country_name[data_cleaning_2$q1_9 == 1] <- 'Luxembourg'
data_cleaning_2$country_name[data_cleaning_2$q1_10 == 1] <- 'Netherlands'
data_cleaning_2$country_name[data_cleaning_2$q1_11 == 1] <- 'Portugal'
data_cleaning_2$country_name[data_cleaning_2$q1_12 == 1] <- 'United Kingdom'
data_cleaning_2$country_name[data_cleaning_2$q1_13 == 1] <- 'Austria'
data_cleaning_2$country_name[data_cleaning_2$q1_14 == 1] <- 'Sweden'
data_cleaning_2$country_name[data_cleaning_2$q1_15 == 1] <- 'Finland'
data_cleaning_2$country_name[data_cleaning_2$q1_16 == 1] <- 'Cyprus'
data_cleaning_2$country_name[data_cleaning_2$q1_17 == 1] <- 'Czech Republic'
data_cleaning_2$country_name[data_cleaning_2$q1_18 == 1] <- 'Estonia'
data_cleaning_2$country_name[data_cleaning_2$q1_19 == 1] <- 'Hungary'
data_cleaning_2$country_name[data_cleaning_2$q1_20 == 1] <- 'Latvia'
data_cleaning_2$country_name[data_cleaning_2$q1_21 == 1] <- 'Lithuania'
data_cleaning_2$country_name[data_cleaning_2$q1_22 == 1] <- 'Malta'
data_cleaning_2$country_name[data_cleaning_2$q1_23 == 1] <- 'Poland'
data_cleaning_2$country_name[data_cleaning_2$q1_24 == 1] <- 'Slovakia'
data_cleaning_2$country_name[data_cleaning_2$q1_25 == 1] <- 'Slovenia'
data_cleaning_2$country_name[data_cleaning_2$q1_26 == 1] <- 'Bulgaria'
data_cleaning_2$country_name[data_cleaning_2$q1_27 == 1] <- 'Romania'
data_cleaning_2$country_name[data_cleaning_2$q1_28 == 1] <- 'Croatia'
data_cleaning_2$country_name[data_cleaning_2$q1_29 == 1] <- 'Other countries'
data_cleaning_2$country_name[data_cleaning_2$q1_30 == 1] <- 'DK'
```

```{r}
# factor
data_cleaning_2$country_name <- factor(data_cleaning_2$country_name)

```

### Extra Country Level Variables

**Legal variable**

We believe that the data was missing some variables related to the legality aspect in a country level so we found this data.

Citation: Williamson, Myles, 2023, "Replication Data for: A Global Analysis of Transgender Rights: Introducing the Trans Rights Indicator Project (TRIP)", <https://doi.org/10.7910/DVN/FXXLTS,> Harvard Dataverse, V1, UNF:6:E4O2rb6DFd44VSZjm9SQNQ== \[fileUNF\]

```{r}
Data_Extra_by_Country <- read_delim("data_extra_by_Country.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)

all_na_columns <- sapply(Data_Extra_by_Country, function(x) all(is.na(x)))

# There are columns with all NA due to the filtering in stata done before for the year 2019.

columns_to_remove <- names(Data_Extra_by_Country)[all_na_columns]

# remove the columns
Data_Extra_by_Country <- Data_Extra_by_Country[, !all_na_columns]
Data_Extra_by_Country <- Data_Extra_by_Country[, !names(Data_Extra_by_Country) %in% "e_polity2"]

# dataframe of all the variables for 2019 related to legal or political views of most of the countries in the world
```

Looking into the extra data set

```{r}
Data_Extra_by_Country$country_name <- factor(Data_Extra_by_Country$country_name)
```

The variable chosen is the the following:

```{r}
Data_Extra_by_Country <- Data_Extra_by_Country %>% 
  select(country_name,adp_general)
```

```{r}
merged_df_1 <- merge(data_cleaning_2, Data_Extra_by_Country, by = "country_name", all.x = TRUE)
```

**Economic Variables**

The economic variables we choose were Unempoyment and GDP per capita.

The data was attained from the World Bank.

Citation: World Bank. (n.d.). DataBank: World Development Indicators \[Economy & Growth\]. Retrieved March 13, 2024, from <https://databank.worldbank.org/source/world-development-indicators>

```{r}
Eco_Extra_Data <- read_delim("Eco Extra Data.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE) 
```

```{r}
merged_df <- merge(merged_df_1, Eco_Extra_Data, by = "country_name", all.x = TRUE)
```

## Descriptive Analysis

### First look at association

-   the descriptive analysis will oscillate between individual-level and country-level analysis

### Question

-   what to do with refusal ? = NA
-   what to do with don't know? = NA

### Gender

Assumption from literature –\> men less supportive than women

```{r}

df_descriptive <- merged_df

df_descriptive$qc19_multinominal <- factor(df_descriptive$qc19, levels = c(1, 2, 3),labels = c("Yes", "No", "Don't know"))

qc19labels <- c("Yes", "No", "Don't know")
names(qc19labels) <- c(1, 2, 3)

df_descriptive |>
  mutate(d10 = as.factor(d10),
         qc19_n = as.numeric(qc19),
         qc19 = as.factor(qc19)) |> 
  group_by(d10) |> 
  mutate(gsumqc19 = sum(qc19_n)) |> 
  ungroup() |> 
  group_by(qc19, d10) |> 
  mutate(gshare19 = (sum(qc19_n)/gsumqc19)*100) |> 
  ungroup() |> 
  ggplot(aes(x = qc19_multinominal, y = gshare19, fill = d10))+
  geom_bar(stat = "identity", position = "dodge")+
  geom_text((aes(label = sprintf("%.1f%%", gshare19))), position = position_dodge(width = 0.9), vjust = -0.3, size = 3) +
  labs(title = "Do you think that transgender persons should be able to change\ntheir civil documents to match their inner gender identity?",
y = NULL,
x = NULL,
caption = "Source: Eurobarometer 91.4") +
  scale_fill_manual(labels = c("Men", "Women"), values = c("blue", "red")) +
  theme_minimal() +
  theme(legend.position = "bottom") +  # Adjust legend position
  guides(fill = guide_legend(title = NULL))  # Hide legend title


```

### Binary conception of gender

predictor of support of US sample

operationalization using question 20 on if there should be a third gender option

Table

```{r}
df_descriptive$qc20_multinominal <- factor(df_descriptive$qc20, levels = c(1, 2, 3),labels = c("Yes", "No", "Don't know"))

CrossTable(df_descriptive$qc19_multinominal, df_descriptive$qc20_multinominal, 
           digits=2, 
           expected=F, 
           asresid=T, 
           chisq=TRUE, 
           prop.chisq=F, 
           format="SPSS")
```

Since the p-value is much smaller than the conventional significance level of 0.05, we reject the null hypothesis. This suggests that there is a significant association between the variables **`qc19`** and **`qc20`**.

Very high convergence of people against a third gender option and rejecting

88.29% of those against a third gender option also negate the option for transgender persons to change their gender in civil documents.

```{r}
plot_binary_1 <-
  df_descriptive |> 
  group_by(country_name) |> 
  summarise(share_no = mean(qc20 == 2, na.rm = TRUE)*100) |> 
  arrange(desc(share_no)) |> 
  mutate(country_name = factor(country_name, levels = country_name)) |>
  filter(country_name != "Otros países") |> 
ggplot(aes(x = country_name, y = share_no)) +
  geom_col(fill = "red") +
  labs(x = "",
       y = "",
       caption = "Source: Eurobarometer 91.4") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.margin = margin(0, 0, 0, 0))  +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  geom_text((aes(label = sprintf("%.1f%%", share_no))),vjust = -0.5, size = 2)  +
  annotate("text", x = 9.5, y = 70, 
           label = "Share of 'No' responses to qc19:\nDo you think that transgender persons should be able to change\ntheir civil documents to match their inner gender identity?",
           hjust = 0, vjust = 0.5, size = 3, color = "black") +
  geom_rect(aes(xmin = 9.3, xmax = 24.3, 
                ymin = 62, ymax = 77),
                fill = "transparent", color = "black", linewidth = 1)

plot_binary_2 <- 
  df_descriptive |> 
  group_by(country_name) |> 
  summarise(share_no = mean(qc19 == 2, na.rm = TRUE)*100) |> 
  arrange(desc(share_no)) |> 
  mutate(country_name = factor(country_name, levels = country_name)) |>
  filter(country_name != "Otros países") |> 
ggplot(aes(x = country_name, y = share_no)) +
  geom_col(fill = "red") +
  labs(x = "",
       y = "",
       caption = "Source: Eurobarometer 91.4") +
  theme_minimal()+ 
  theme(axis.text.x = element_text(angle = 90, hjust= 0),
        plot.margin = margin(-10, 0, 0, 0)) +
  scale_x_discrete(position = "top") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  geom_text((aes(label = sprintf("%.1f%%", share_no))), vjust = -0.5, size = 2)  +
  annotate("text", x = 10, y = 65, 
           label = "Share of 'No' responses to qc20:\nDo you believe that official documents, like passports and birth\ncertificates, should have a third option, such as X or O (other)\nbeside male (M) and female (F)for those persons who do not\nidentify as female and male?",
           hjust = 0, vjust = 0.5, size = 3, color = "black") +
  geom_rect(aes(xmin = 9.8, xmax = 24.6, 
                ymin = 53, ymax = 76),
                fill = "transparent", color = "black", linewidth = 1)
```

```{r}
#| echo: false
#| layout-ncol: 1

plot_binary_1
plot_binary_2
```

-   visually this high convergence can also be seen on the country level

-   to verify this statistically, the Spearman rank test is used as the country-level proportions are not normally distributed

-   the statistical association between agreeing there should be a third gender option and supporting the ability for transgender people to change their civil document on the country level is estimated

-   the Spearman rank test reports a very strong positive relationship between the two variables supporting the previous assertion that the acceptance of a third gender differentiates countries concerning their average approval of the civil document alteration for transgender people

```{r}
df_country_1 <- 
  df_descriptive |>
  group_by(country_name) |> 
  summarize(
    prop_gndr_bin = (sum(qc20_multinominal == "Yes") / n())*100, # Calculate proportion with high education
    prop_qc19_yes = (sum(qc19 == 1) / n())*100)


cor(df_country_1$prop_gndr_bin, df_country_1$prop_qc19_yes, method = "spearman")
```

### Political ideology

```{r}
df_descriptive$d1_ordinal <- factor(df_descriptive$d1, levels = c(1, 2, 3, 4 , 5, 6, 7, 8, 9, 10, 97, 98),labels = c("Left", "2", "3", "4", "5", "6", "7", "8", "9", "Right", "Non-response", "Don't know"))

CrossTable(df_descriptive$d1_ordinal, df_descriptive$qc19_multinominal,
           digits=2, 
           expected=F, 
           asresid=T, 
           chisq=TRUE, 
           prop.chisq=F, 
           format="SPSS")
```

Note:

-   25% of people chose 5, the exact middle ground

Surprising results:

-   The inherent idea is upheld that a greater share of more left-leaning people are in favor

-   However, difference to very right-leaning people is not that great (60-30 left vs. 43-45 right)

Problem:

-   self-reporting self-right scale might not be so accurate

-   a lot of information is hidden in the non-response and don't know groups of political ideology as they have greater don't know rates for qc19

Take-away

-   Still significant association (but maybe not as strong as suspected)

```{r}

df_descriptive |> 
  mutate(d1_bin = case_when(
    d1 < 4 ~ "Left",
    d1 < 8 ~ "Middle",
    d1 < 11 ~ "Right",
    TRUE ~ "DK/Refusal"
  )) |> 
  mutate(d1_bin = fct_relevel(d1_bin, "DK/Refusal", after = Inf)) |> 
  group_by(d1_bin, qc19_multinominal) |> 
  summarise(count = n()) |> 
  group_by(d1_bin) |> 
  mutate(share = count / sum(count) * 100) |>
  ungroup() |> 
  ggplot(aes(x = d1_bin, y = share, fill = qc19_multinominal))+
  geom_bar(stat = "identity", position = "dodge")+
  geom_text((aes(label = sprintf("%.1f%%", share))), position = position_dodge(width = 0.9), vjust = -0.3, size = 3) +
  theme_minimal() +
  labs(title = "Support of change of gender in civil documents for transgender\npeople by political ideology",
       x = "Political Ideology",
       y = "",
       caption = "Source: Eurobarometer 91.4")+
  theme(legend.position = "bottom",
        legend.title = element_blank()) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  scale_fill_brewer(palette = "Dark2")
```

#### Add government ideology as better proxy

### Religiosity

```{r}
df_descriptive <-
  df_descriptive |> 
  mutate(sd3_cat = case_when(sd3 == 1 ~ "Catholic",
                             sd3 == 2 ~ "Orthodox Christian",
                             sd3 == 3 ~ "Protestant",
                             sd3 == 4 ~ "Other Christian",
                             sd3 == 5 ~ "Jewish",
                             sd3 < 9 ~ "Muslim", 
                             sd3 < 12 ~ "Other",
                             sd3 < 14 ~ "Atheist | Agnostic",
                             sd3 == 14 ~ "Other",
                             sd3 > 14 ~ "Don't know | Refusal")) 

CrossTable(df_descriptive$sd3_cat, df_descriptive$qc19_multinominal,
           digits=2, 
           expected=F, 
           asresid=T, 
           chisq=TRUE, 
           prop.chisq=F, 
           format="SPSS")
```

Problem:

-   Very low counts for Jewish, Muslim, Other and Other Christian as can be seen in the plot below

```{r}
df_descriptive |> 
  mutate(sd3_cat = fct_relevel(sd3_cat, "Don't know | Refusal", after = Inf)) |> 
  ggplot(aes(x = sd3_cat)) +
  geom_bar(fill = "black", color = "black", alpha = 0.5) +
  labs(title = "Frequency by religious affiliation",
       x = "",
       y = "Absolutec count",
       caption = "Source: Eurobarometer 91.4") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Findings:

-   on the individual level not believing in God more say yes as well as Protestant

-   Catholics in favor but less strongly

-   Orthodox Christian strong outlier with being against it

Summary:

-   roughly one may say that countries with a higher share of Orthodox Christians may be rather against it

```{r}
df_descriptive |> 
  group_by(country_name) |> 
  summarise(share = mean(sd3_cat == "Orthodox Christian", na.rm = TRUE)*100) |> 
  arrange(desc(share)) |> 
  mutate(country_name = factor(country_name, levels = country_name)) |>
  filter(country_name != "Otros países") |> 
ggplot(aes(x = country_name, y = share)) +
  geom_col(fill = "red") +
  labs(title = "Share of Orthodox Christians", 
       x = "",
       y = "",
       caption = "Source: Eurobarometer 91.4") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.margin = margin(0, 0, 0, 0))  +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  geom_text((aes(label = sprintf("%.1f%%", share))),vjust = -0.5, size = 2)
```

### Contact

```{r}
df_descriptive <-
  df_descriptive |> 
  mutate(
    sd1_7_NA = car::recode(sd1_7, "1=1; 2=0; 3=NA; 4=2"),
    sd1_7_multinomial = factor(sd1_7_NA, levels = c(0, 1, 2), labels = c("No", "Yes", "Don't know"))
  )

CrossTable(df_descriptive$sd1_7_multinomial, df_descriptive$qc19_multinominal,
           digits=2, 
           expected=F, 
           asresid=T, 
           chisq=TRUE, 
           prop.chisq=F, 
           format="SPSS")
```

-   association is statistically significant

-   the majority of people did not have contact with a transgender person

-   there is a clear tendency on the individual level that among those that have not had contact with a transgender person the approval is relatively lower as for those who have had the margin is very clear (76% vs 18%)

-   contact is associated rather with approval. Not having had contact not associated with being against, just lower approval

-   this is visualised below

```{r}
df_descriptive|> 
  drop_na(sd1_7_multinomial) |> 
  group_by(sd1_7_multinomial, qc19_multinominal) |> 
  summarise(count = n()) |> 
  ungroup() |> 
  group_by(sd1_7_multinomial) |> 
  mutate(share = count / sum(count) * 100) |>
  ungroup() |> 
  ggplot(aes(x = sd1_7_multinomial, y = share, fill = qc19_multinominal))+
  geom_bar(stat = "identity", position = "dodge")+
  geom_text((aes(label = sprintf("%.1f%%", share))), position = position_dodge(width = 0.9), vjust = -0.3, size = 3) +
  theme_minimal() +
  labs(title = "Support of change of gender in civil documents for transgender\npeople by previous conatct with transgender person",
       x = "Contact",
       y = "",
       caption = "Source: Eurobarometer 91.4")+
  theme(legend.position = "bottom",
        legend.title = element_blank()) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  scale_fill_brewer(palette = "Dark2")
```

### Age

-   using anova to seek out differences between those who responded 'yes', 'no', don't know' by age

-   p-value let's us reject the null hypothesis. There is a statistically significant difference between at least two groups

```{r}
anova <- aov(d11 ~ qc19_multinominal, data = df_descriptive)
summary(anova)
```

-   visualizing group difference using boxplots

-   showing that the strongest difference is between those saying 'yes' and those stating 'don't know'

-   implies there is valuable information from particularly older people missing

```{r}
ggplot(df_descriptive, aes(x = qc19_multinominal, y = d11, fill = "red")) +
  geom_boxplot(color = "red") +
  theme_minimal() +
  labs(title = "Support of change of gender in civil documents for transgender\npeople by age",
       y = "Age",
       x = "Support",
       caption = "Source: Eurobarometer 91.4") +
  theme(legend.position = "none")
```

-   regarding the difference between 'yes' and 'no' - those reporting the former are generally a bit younger

-   employ a correlation to look at strength and direction of the association while differentiating only between Yes and No

```{r}
df_cor_age <- 
  df_descriptive |> 
  select(qc19, d11) |> 
  filter(qc19 < 3) |> # filtering out everything besides 1 and 2
  mutate(qc19 = ifelse(qc19 == 2, 0, qc19)) # Assigning 2 the number 0 for logical interpretation

cor(df_cor_age$d11, df_cor_age$qc19)
```

-   as already implied above only a slight relation ship

-   negative direction shows that a one year increase in age decreases approval

### Awareness

-   if type of discrimination is widespread in own country

```{r}
df_descriptive$qc1_8_ordinal <- factor(df_descriptive$qc1_8, levels = c(1, 2, 3, 4 , 5, 6),labels = c("Very wide spread", "Fairly wide spread", "Fairly rare", "Very rare", "Nonexistent
", "DK"))

CrossTable(df_descriptive$qc1_8_ordinal, df_descriptive$qc19_multinominal,
           digits=2, 
           expected=F, 
           asresid=T, 
           chisq=TRUE, 
           prop.chisq=F, 
           format="SPSS")
```

-   p-value lets us very confidently state that there is a statistically significant relationship between the variables

-   among those reporting transgender people face very or fairly wide spread discrimination, the relative support for the possibility to change civil documents is greater than for those stating it is fairly rare, while among those reporting it is very rare the share between supporters and opponents is balanced

-   this clear association between more awareness and greater support is visible below

```{r}
df_descriptive|> 
  drop_na(qc1_8_ordinal) |> 
  group_by(qc1_8_ordinal, qc19_multinominal) |> 
  summarise(count = n()) |> 
  ungroup() |> 
  group_by(qc1_8_ordinal) |> 
  mutate(share = count / sum(count) * 100) |>
  ungroup() |> 
  ggplot(aes(x = qc1_8_ordinal, y = share, fill = qc19_multinominal))+
  geom_bar(stat = "identity", position = "dodge")+
  geom_text((aes(label = sprintf("%.1f%%", share))), position = position_dodge(width = 0.9), vjust = -0.3, size = 3) +
  theme_minimal() +
  labs(title = "Support of change of gender in civil documents for transgender\npeople by perceived discrimination transgender people face",
       x = "Contact",
       y = "",
       caption = "Source: Eurobarometer 91.4")+
  theme(legend.position = "bottom",
        legend.title = element_blank()) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  scale_fill_brewer(palette = "Dark2")
```

-   between countries

-   countries which have a higher share of approval are to be found rather towards the

```{r}
df_descriptive |> 
  mutate(qc1_8_ordinal = case_when(
    qc1_8_ordinal == "Very wide spread" ~ "Very or fairly wide spread",
    qc1_8_ordinal == "Fairly wide spread" ~ "Very or fairly wide spread",
    TRUE ~ qc1_8_ordinal
  )) |>
  group_by(country_name) |> 
  summarise(share = mean(qc1_8_ordinal == "Very or fairly wide spread", na.rm = TRUE)*100) |> 
  arrange(desc(share)) |> 
  mutate(country_name = factor(country_name, levels = country_name)) |>
  filter(country_name != "Otros países") |> 
ggplot(aes(x = country_name, y = share)) +
  geom_col(fill = "red") +
  labs(title = "Share of respondents perceiving discrimination against transgender people\nto be very of fairly wide spread", 
       x = "",
       y = "",
       caption = "Source: Eurobarometer 91.4") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.margin = margin(0, 0, 0, 0))  +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  geom_text((aes(label = sprintf("%.1f%%", share))),vjust = -0.5, size = 2)
```

-   to further investigate the country-level differences the association of the proportion of individuals aware of discrimination per country and the proportion of individuals in favor of the ability of transgender people to change their civil documents

-   because the variables are non-normally distributed, we use the Spearman rank test

```{r}
df_country_2 <-
  df_descriptive |> 
  mutate(qc1_8_ordinal = case_when(
    qc1_8_ordinal == "Very wide spread" ~ "Very or fairly wide spread",
    qc1_8_ordinal == "Fairly wide spread" ~ "Very or fairly wide spread",
    TRUE ~ qc1_8_ordinal
  )) |> 
  group_by(country_name) |> 
  summarize(
    prop_dis_wide = (sum(qc1_8_ordinal == "Very or fairly wide spread") / n())*100, # Calculate proportion with high education
    prop_qc19_yes = (sum(qc19 == 1) / n())*100)


cor(df_country_2$prop_dis_wide, df_country_2$prop_qc19_yes, method = "spearman")
```

-   the test shows a strong relationship between awareness of discrimination faced by transgender people on the country-level insuating that in countries with greater awareness there is greater support

### Solidarity among minority groups

```{r}
df_descriptive <- 
  df_descriptive |>  
  mutate(
    sd2t = factor(sd2t, levels = c(0, 1), labels = c("Not mentioned", "Yes"))
  )
```

As the subsequent plot demonstrates, there is no general solidarity among minority group expressed by a greater support for transgender people than among individuals without any minority-background.

```{r}
df_descriptive|> 
  drop_na(sd2t) |> 
  group_by(sd2t, qc19_multinominal) |> 
  summarise(count = n()) |> 
  ungroup() |> 
  group_by(sd2t) |> 
  mutate(share = count / sum(count) * 100) |>
  ungroup() |> 
  ggplot(aes(x = sd2t, y = share, fill = qc19_multinominal))+
  geom_bar(stat = "identity", position = "dodge")+
  geom_text((aes(label = sprintf("%.1f%%", share))), position = position_dodge(width = 0.9), vjust = -0.3, size = 3) +
  theme_minimal() +
  labs(title = "Support of change of gender in civil documents for transgender\npeople by perceived discrimination transgender people face",
       x = "Contact",
       y = "",
       caption = "Source: Eurobarometer 91.4")+
  theme(legend.position = "bottom",
        legend.title = element_blank()) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  scale_fill_brewer(palette = "Dark2")
```

However, this support is much greater among people who themselves are a sexual minority. As such, people with similar experiences as transgender people are more likely to support transgender rights.

```{r}
df_descriptive <- 
  df_descriptive |>  
  mutate(
    sd2_5 = factor(sd2_5, levels = c(0, 1), labels = c("Not mentioned", "Yes"))
  )


df_descriptive|> 
  drop_na(sd2_5) |> 
  group_by(sd2_5, qc19_multinominal) |> 
  summarise(count = n()) |> 
  ungroup() |> 
  group_by(sd2_5) |> 
  mutate(share = count / sum(count) * 100) |>
  ungroup() |> 
  ggplot(aes(x = sd2_5, y = share, fill = qc19_multinominal))+
  geom_bar(stat = "identity", position = "dodge")+
  geom_text((aes(label = sprintf("%.1f%%", share))), position = position_dodge(width = 0.9), vjust = -0.3, size = 3) +
  theme_minimal() +
  labs(title = "Support of change of gender in civil documents for transgender\npeople by perceived discrimination transgender people face",
       x = "Contact",
       y = "",
       caption = "Source: Eurobarometer 91.4")+
  theme(legend.position = "bottom",
        legend.title = element_blank()) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  scale_fill_brewer(palette = "Dark2")
```

### Education

-   age when leaving formal education

-   expected that higher and lower ages are outliers

-   one is able to gauge differences between the 'yes' and 'no' respondents with the former reporting a slightly higher age as their mode (highest frequency)

```{r}
df_descriptive <- 
  df_descriptive |> 
  mutate(d8 = case_when(d8 < 98 ~ d8,
                        TRUE ~ NA)) # turning don't know and refusal into NA to keep continuous character of the variable - imputed later


ggplot(df_descriptive, aes(x = qc19_multinominal, y = d8, fill = "red")) +
  geom_boxplot(color = "red") +
  theme_minimal() +
  labs(title = "Support of change of gender in civil documents for transgender\npeople by age",
       y = "Age when leaving formal education",
       x = "Support",
       caption = "Source: Eurobarometer 91.4") +
  theme(legend.position = "none")

```

The ANOVA analysis supports the claim, at least, that the groups are statistically significantly different from one another.

```{r}
anova <- aov(d8 ~ qc19_multinominal, data = df_descriptive)
summary(anova)
```

```{r}
df_cor_age <- 
  df_descriptive |> 
  select(qc19, d8) |> 
  filter(qc19 < 3) |> # filtering out everything besides 1 and 2
  mutate(qc19 = ifelse(qc19 == 2, 0, qc19)) # Assigning 2 the number 0 for logical interpretation

cor(df_cor_age$d8, df_cor_age$qc19)
```

The correlation shows there is a very weak association in favor of the literature's case that being more educated is associated with an increased support of transgender rights.

### The economic situation

-   During the last twelve months, would you say you had difficulties to pay your bills at the end of the month…?

```{r}

df_descriptive <- 
  df_descriptive |> 
  mutate(
    d60 = car::recode(d60, "4=NA"),
    d60_ordinal = factor(d60, levels = c(1, 2, 3), labels = c("Most of the time", "From time to time", "Almost never/ never"))
  )

CrossTable(df_descriptive$d60_ordinal, df_descriptive$qc19_multinominal,
           digits=2, 
           expected=F, 
           asresid=T, 
           chisq=TRUE, 
           prop.chisq=F, 
           format="SPSS")
```

-   following the literature, among those who never or almost never had difficulties to pay their bills at the end of the month, the support for the possibility of transgender people to alter their gender in their civic documents is markedly higher.

-   the low p-value indicates that the association is statistically significant

-   confers with literature

-   this difference is visualized by the graph below

```{r}
df_descriptive|> 
  drop_na(d60_ordinal) |> 
  group_by(d60_ordinal, qc19_multinominal) |> 
  summarise(count = n()) |> 
  ungroup() |> 
  group_by(d60_ordinal) |> 
  mutate(share = count / sum(count) * 100) |>
  ungroup() |> 
  ggplot(aes(x = d60_ordinal, y = share, fill = qc19_multinominal))+
  geom_bar(stat = "identity", position = "dodge")+
  geom_text((aes(label = sprintf("%.1f%%", share))), position = position_dodge(width = 0.9), vjust = -0.3, size = 3) +
  theme_minimal() +
  labs(title = "Support of change of gender in civil documents for transgender\npeople by times having difficulties to pay the bills in last 12 months",
       x = "Having difficulties to pay the bills in last 12 months",
       y = "",
       caption = "Source: Eurobarometer 91.4")+
  theme(legend.position = "bottom",
        legend.title = element_blank()) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  scale_fill_brewer(palette = "Dark2")
```

### Support for LGBT rights

```{r}
df_descriptive <- 
  df_descriptive |> 
  mutate(
    qc15_1_ordinal = factor(qc15_1, levels = c(1, 2, 3, 4, 5), labels = c("Totally agree", "Tend to agree", "Tend to disagree", "Totally disagree", "Don't know"))
  )

CrossTable(df_descriptive$qc15_1_ordinal, df_descriptive$qc19_multinominal,
           digits=2, 
           expected=F, 
           asresid=T, 
           chisq=TRUE, 
           prop.chisq=F, 
           format="SPSS")
```

-   the p-value indicates statistical significance concerning the association between the two variables

-   as may be expected, among those disagreeing that people part of the LGB community should have equal rights the support is very low and the rejection of the possibility

-   rejection is almost as pronounced as the support regarding transgender people's ability to change their gender identification of those who totally agree people of the LGB community should have the same rights

-   important: two way association

-   on the country level

```{r}
df_country_3 <- 
  df_descriptive |>
  mutate(qc15_1_ordinal = case_when(
    qc15_1_ordinal == "Totally agree" ~ "Totally or tend to agree",
    qc15_1_ordinal == "Tend to agree" ~ "Totally or tend to agree",
    TRUE ~ qc15_1_ordinal
  )) |> 
  group_by(country_name) |> 
  summarize(
    prop_lgb_yes = (sum(qc15_1_ordinal == "Totally or tend to agree") / n())*100, # Calculate proportion with high education
    prop_qc19_yes = (sum(qc19 == 1) / n())*100)


cor(df_country_3$prop_lgb_yes, df_country_3$prop_qc19_yes, method = "spearman")
```

-   very strong association on the country level indicating that most likely countries where there is greater support for the legal equality for LGB people also post higher support rates for the adoption of legislation allowing transgender people to change their gender in civic documents

-   this only captures the support direction. As outlined above, a similar strength may be expected for the revers association.

### Legal protections for the transgender community in 2019 by country level

To observe if there is are general legal protections across various countries within our dataset, an interactive map was developed.

```{r}
df <- df_descriptive %>% 
  select(country_name, adp_general)
df$country_name <- as.character(df$country_name)

df_unique <- df %>%
  group_by(country_name) %>%
  summarise(adp_general = max(adp_general)) %>%
  ungroup()

europe_map <- ne_countries(scale = "medium", continent = "europe", returnclass = "sf")

df_joined <- left_join(europe_map, df_unique, by = c("sovereignt" = "country_name"))

```

```{r}
df_joined$status <- case_when(
  df_joined$adp_general == 1 ~ "General",
  df_joined$adp_general == 0 ~ "None",
  TRUE ~ "No Information"
)

df_plot <- df_joined %>% 
  select(sovereignt, status)

df_plot <- na.omit(df_plot)

```

```{r}
pal <- colorFactor(c("#2ca02c", "#377eb8", "#d62728"), domain = df_plot$status)

plot1 <- leaflet(df_plot) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(fillColor = ~pal(status),
              color = "black",
              fillOpacity = 0.7,
              weight = 1,
              popup = ~paste(sovereignt, status)) %>%
  addLegend(pal = pal, values = ~status,
            title = "Status",
            position = "bottomright")
plot1
```

**Association**

```{r}
table <- table(df_descriptive$qc19_multinominal,df_descriptive$adp_general)
chisq.test(table)

```

With this results we can conclude that there is a statistically significant relationship between the our dependent variables and the variable that establish if a country has general legal protections against discrimination for transgender individuals.

Descriptive analysis for GDP

Descriptive analysis for Unemployment

## Modeling

**Dependent Variable (qc19)**

It relates to this question:

*Do you think that tran*sgender persons should be able to change their civil documents to match their inner gender identity?

It takes the values:

-   1 for "Yes"

-   2 for "No"

We will take the variable as Binary in the individual level

```{r}
df_descriptive <- df_descriptive %>%
  mutate(binary_qc19 = ifelse(qc19 == 1, 1, 
                              ifelse(qc19 == 2, 0, NA)))
# there are 3280 NA's

Individual_data <- df_descriptive %>%
  select(serialid,country_name, binary_qc19)
```

INDIVIDUAL LEVEL

**Gender**

```{r}
df_descriptive <- df_descriptive %>%
  mutate(
    male = car::recode(d10, "2=0; 1=1")
  )

Individual_data <- left_join(Individual_data, df_descriptive %>% select(serialid, male), by = "serialid")
```

**Age**

The different categories for age will be transformed as factor for better analysis. There are 3 variables that represent four, six and seven categories of age.

```{r}
df_descriptive <- df_descriptive %>%
  mutate(
    Cat_age_four = factor(d11r1, 
                           labels = c("15 - 24", "25 - 39", "40 - 54", ">=55")),
    Cat_age_six = factor(d11r2, 
                           labels = c("15 - 24", "25 - 34", "35 - 44", "45 - 54", "55 - 64", ">=65")),
    Cat_age_seven = factor(d11r3, 
                           labels = c("15 - 24", "25 - 34", "35 - 44", "45 - 54", "55 - 64", "65 - 74", ">=75"))
  )


Individual_data <- left_join(Individual_data, df_descriptive %>% select(serialid, Cat_age_four,Cat_age_six,Cat_age_seven), by = "serialid")
```

```{r}
Individual_data <- left_join(Individual_data, df_descriptive %>% select(serialid, d11), by = "serialid")
```

**Political_ideology**

In political matters people talk of "the left" and "the right". How would you place your views on this scale?

```{r}

df_descriptive$d1_ordinal <- relevel(df_descriptive$d1_ordinal, ref = "Left") 
df_descriptive$d1_ordinal[df_descriptive$d1_ordinal %in% c("Non-response", "Don't know", 98)] <- NA

df_descriptive$d1_ordinal <- factor(df_descriptive$d1_ordinal)

df_descriptive <- df_descriptive %>% 
  rename(political_ideology = d1_ordinal) 

Individual_data <- left_join(Individual_data, df_descriptive %>% select(serialid, political_ideology), by = "serialid")

```

**Religion_cat**

Religion - Do you consider yourself to be…?

```{r}

df_descriptive <- df_descriptive %>%
  mutate(Religion_cat = case_when(
    sd3_cat == "Catholic" ~ "Catholic",
    sd3_cat == "Orthodox Christian" ~ "Orthodox Christian",
    sd3_cat == "Protestant" ~ "Protestant",
    sd3_cat == "Other Christian" ~ "Other Christian",
    sd3_cat == "Jewish" ~ "Jewish",
    sd3_cat == "Muslim" ~ "Muslim", 
    sd3_cat == "Other" ~ "Other Religion",
    sd3_cat == "Atheist | Agnostic" ~ "Atheist | Agnostic",
    sd3_cat == "Other" ~ "Other Belief",
    TRUE ~ "Other"
  ))

Individual_data <- left_join(Individual_data, df_descriptive %>% select(serialid, Religion_cat), by = "serialid")

```

**sd1_7**

Do you have friends or acquaintances who are transgender?

```{r}

df_descriptive <- df_descriptive %>%
  mutate(sd1_7_factor = factor(sd1_7, 
              levels = c(1, 2, 3, 4),
              labels = c("Yes", "No", "Refusal (SPONTANEOUS)", "Don't know")))

Individual_data <- left_join(Individual_data, df_descriptive %>% select(serialid, sd1_7_factor), by = "serialid")
```

**d8**

How old were you when you stopped full‐time education?

```{r}

df_descriptive <- df_descriptive %>%
  mutate(age_stopped_education = case_when(
    d8 == 98 ~ NA_real_,  # Refusal treated as missing
    d8 == 99 ~ NA_real_,  # Don't know treated as missing
    d8 == 00 ~ NA_real_,  # Still Studying can be treated as missing or given a special code
    d8 == 01 ~ NA_real_,  # No Education can be treated as missing or given a special code
    TRUE ~ as.numeric(d8) # Convert all other responses to numeric
  ))



Individual_data <- left_join(Individual_data, df_descriptive %>% select(serialid, age_stopped_education), by = "serialid")
```

**d60_ordinal** During the last twelve months, would you say you had difficulties to pay your bills at the end of the month…?

```{r}
Individual_data <- left_join(Individual_data, df_descriptive %>% select(serialid, d60_ordinal), by = "serialid")

```

**Sexual minority(sd2_5)**

Where you live, do you consider yourself to be part of a sexual minority?

```{r}
Individual_data <- left_join(Individual_data, df_descriptive %>% select(serialid, sd2_5), by = "serialid")
```

**d25**

Would you say you live in a...?

```{r}
df_descriptive <- df_descriptive %>%
  mutate(area_type = factor(d25, 
                                levels = c(1, 2, 3, 4),
                                labels = c("Rural area or village", 
                                           "Small or middle-sized town", 
                                           "Large town", 
                                           "Don't know")))


Individual_data <- left_join(Individual_data, df_descriptive %>% select(serialid, area_type), by = "serialid")

```

**qc15_1_ordinal**

```{r}
Individual_data <- left_join(Individual_data, df_descriptive %>% select(serialid, qc15_1_ordinal), by = "serialid")

```

**qc1_8_ordinal**

```{r}
Individual_data <- left_join(Individual_data, df_descriptive %>% select(serialid, qc1_8_ordinal), by = "serialid")
```

COUNTRY LEVEL

**Variables created earlier country level.**

Examining the influence of country-level proportion of respondents to a particular outcome. Aimed at understanding differences between countries on various outcomes.

```{r}
Country_level <- reduce(list(df_country_1 %>% select(-prop_qc19_yes), 
                             df_country_2 %>% select(-prop_qc19_yes), 
                             df_country_3 %>% select(-prop_qc19_yes)), 
                        full_join, by = "country_name")


Country_level <- left_join(Country_level, binary_qc19_data %>% select(country_name, count_yes,count_no),  by = "country_name")

```

**adp_general**

A country-level variable indicating general legal protections against discrimination for transgender individuals, with 0 representing "No" (no protections) and 1 representing "Yes" (protections exist)

```{r}
Country_level <- left_join(Country_level, df_descriptive %>% 
  select(country_name, trans_legal_protection = adp_general) %>% 
  distinct(), by = "country_name")

```

**Country_name**

To control for country effects or to specifically examine how outcomes vary by country we will include the variable created earlier, "country_name".

Unemployment, total (% of total labor force)

```{r}
Country_level <- left_join(Country_level, df_descriptive %>% 
  select(country_name, Unemployment = `Unemployment. total (% of total labor force) (modeled ILO estimate)`) %>% 
  distinct(), by = "country_name")

```

GDP per capita (current US\$)

```{r}
Country_level <- left_join(Country_level, df_descriptive %>% 
  select(country_name, Gdp_per_capita = `GDP per capita (current US$)`) %>% 
  distinct(), by = "country_name")

```

DATA FINAL COUNTRY

```{r}
Data_final_Cntry <- Country_level |>
   select(prop_gndr_bin, prop_dis_wide,prop_lgb_yes, trans_legal_protection, Gdp_per_capita, Unemployment)
```

DATA FINAL INDIVIDUAL

```{r}

Data_final_Ind <- Individual_data |> 
  select(serialid,country_name,binary_qc19,male, d11, Cat_age_four,Cat_age_six, Cat_age_seven,political_ideology,Religion_cat , sd1_7_factor, age_stopped_education, d60_ordinal,sd2_5,area_type,qc15_1_ordinal)

```

### Regression Individual Level

Imputation

Regression

```{r}
data_reg_ind <- Data_final_Ind |> 
  select(binary_qc19,male, d11 ,political_ideology,Religion_cat , sd1_7_factor, age_stopped_education, d60_ordinal,sd2_5,area_type,qc15_1_ordinal)


data_reg_ind = data_reg_ind %>%
  filter(!is.na(binary_qc19))
```

```{r}

logit_model <- glm(binary_qc19 ~ ., family=binomial(link='logit'), data= data_reg_ind)

summary(logit_model)
```

There are several variables that are not significant. We also tried the regression with the categories of variables for age that we had and they were also not significant. We will now proceed to reduce the variables used.

```{r}
logit_model2 <- glm(binary_qc19 ~ male + d11 + political_ideology +                         Religion_cat + sd1_7_factor + d60_ordinal + qc15_1_ordinal, 
                    family = binomial(link = "logit"), data = data_reg_ind)

summary(logit_model2)


```

We are going to transform the age variable to include a quadratic term as we suspect the relationship between age and the dependent variable is not linear.

```{r}
logit_model3 <- glm(binary_qc19 ~ male + d11 + I(d11^2)  + political_ideology +                         Religion_cat + sd1_7_factor + d60_ordinal + qc15_1_ordinal, 
                    family = binomial(link = "logit"), data = data_reg_ind)

summary(logit_model3)
```

**Interpretation**

```{r}
exp(coef(logit_model3))
```

**Male:** The odds of supporting Transgender individuals of being able to change their civil document to match their inner gender are about 25.4% lower for males compared to females, since the odds ratio is 0.74.

**d11:** For each one-year increase in age, the odds of the being supportive of Transgender individuals changing their civil document increase by about 1.9%.

**political_ideology:** Identifying with the political ideology "Right" decreases the odds of the supporting the change in civil documents by transgender individuals to match their inner gender identity by about 22.8%, compared to the baseline category of political ideology.

**Religion_cat:** Identifying as having Catholics faith have a decrease in the odds of about 21.2%, while identifying as having Jewish faith have an odds increase of about 2.3%, in comparison to the baseline category religion.

**sd1_7_factor:** Responding "No" to having friends or acquaintances who are transgender decreases the odds of the supporting Transgender individuals in being able to change their civil documents by about 45.5%, in comparison to the baseline category.

**d60_ordinal:** Responding that during the last twelve month you had difficulties from time to time in paying your bills, slightly increases the odds of being in favor of letting transgender individuals to change their civil document by about 7.5%, while "Almost never/never" increases the odds by about 26.3%.

**qc15_1_ordinal:** Tending to agree that Gay, lesbian and bisexual people should have the same rights as heterosexual people, decreases the odds by about 61.5%, while "Totally disagree" decreases the odds by about 92.9%.

### Regression for Individual and Country Level Analysis. 

## Prediction by Individual Level

We want to create a model that is able to predict the support for transgender individuals changing civil documents.

### Definition of support

```{r}
prop_table_yes_no <- prop.table(table(factor(data_reg_ind$binary_qc19, levels = c(0, 1), labels = c("No", "Yes"))))

prop_table_yes_no
```

We need to find the factors that are causing this 60% of support for our dependent variable..

### Data splitting

```{r}

set.seed(123)

in_train <- createDataPartition(y = data_reg_ind$binary_qc19, p = 0.8, list = FALSE)

in_train <- in_train[, 1]

training <- data_reg_ind[in_train, ]
testing <- data_reg_ind[-in_train, ]
```

//////////////////////////

```{r}
 
# Check for NA values in the data
na_summary <- colSums(is.na(complete_data))
 
# Print columns with NA values
cols_with_na_names <- names(na_summary[na_summary > 0])
cols_with_na <- na_summary[na_summary > 0]
print(cols_with_na)

```

impute using fixed effects, use a vector to include the factor country

multilevel analysis on the countries, join with country level data

IGNORE THIS ! imputation code

```{r}
library(mice)
library(missForest)

# Define the number of imputations
num_imputations <- 5  # Adjust as needed

# Perform multiple imputation using polyreg method
imputed_data <- mice(complete_data, method = "polyreg", m = num_imputations)

# Get the completed datasets
imputed_datasets <- complete(imputed_data)

# Subset the dataset to include only the variables with warnings
vars_with_warnings <- c("d40_d11", "p7ro", "p6ro", "d15b", "p7", "p7gr_r", "p7ie", "p7gr", "p7pl_r", "p7fi", "p6si", "p6fi", "p7lv", "p6hr", "p6lt", "p7cy", "p7hu_r", "p7pt", "p7sk", "d15b_r", "p6mt", "p7pl", "p7be_r", "p6lv", "p7ro_r", "p7ee", "p6hu", "p7lt", "p7it_r2", "p7it", "p13", "p6sk", "p6cy", "p6de", "p7de", "p7it_r1", "p7es", "p7hu", "p6fr", "p7be", "p7se_r")
subset_data <- complete_data[vars_with_warnings]

# Check the structure of the subsetted dataset
str(subset_data)

# Perform multiple imputation using polyreg method
imputed_data <- mice(subset_data, method = "polyreg", m = num_imputations)

# Get the completed datasets
imputed_datasets <- complete(imputed_data)
```

Alex part of the code

impute something?

```{r}

```

make the binary predictive model using the new merged and reduced data set, try all the models to see which is best

```{r}

```

figure out what to do with DK and refusal
